---
title: "Plots - tool function experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  word_document: default
  html_document:
    df_print: paged
---


```{r Knitr Options, include=FALSE}
	knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)

```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"

```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(ggfortify)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)

```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in

```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)

```

The imported file is: "`r paste0("~/", data_file)`"  


---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 24:length(imp_data)

```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")

```


---

# Plot each of the selected numeric variable 
## (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
# BottleRocket2 
custom.col5 <- data.frame(type = levels(imp_data$Edge.angle), col = c("#046C9A", "#FAD510")) imp_data$col <- custom.col5[imp_data$Edge.angle, "col"]


for (i in num.var){
  #plot
  range_var <- range(imp_data[[i]]) # gets the min/max range of the data set 

	p <- ggplot(data=imp_data, aes_string(x = "Task", y=names(imp_data)[i],  
	                                 colour = "Edge.angle")) + 
	     # avoids overplotting
	     geom_jitter(size = 3,  position = position_jitter(width = 0.4, seed = 1)) + 
	     coord_cartesian(ylim = range_var) + 
	     theme_classic() + 
	     scale_colour_manual(values = custom.col5$col) +
	     # removes the "." between "Edge.angle" in the legend 
	     labs(colour = gsub("\\.", " ", "Edge.angle")) + 
	     # removes the "." between the variable names
	     labs(y = gsub("\\.", " ", names(imp_data)[i])) + 
       # combines the flint and the lydite plot into one    
	     facet_wrap(~Raw.material) 
	print(p)


	#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_plot_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")
}

```


## Principal component analysis (without outliers)
```{r}
# excludes the outliers 
# adds the indices as row numbers 
imp_data <- imp_data %>% mutate(id = row_number())
imp_data2 <- imp_data[-c(2, 10), ]

# removes the rows with missing values 
data_pca <- na.omit(imp_data2) 
imp_data.pca <- prcomp(data_pca[, c(24:25, 42, 47, 51, 54, 56)], scale. = TRUE) 

# FantasticFox1 
custom.col4 <- data.frame(type = levels(data_pca$Task), col = c("#E58601", "#B40F20")) 
data_pca$col <- custom.col4[data_pca$Task, "col"]


# plots the task - PCA with convex hull 
# Using ggfortify
a<- autoplot(imp_data.pca, data = data_pca, colour = "Task", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "convex", frame.colour = "Task", frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col4$col)
         
print(a)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_Task", ".pdf")
ggsave(filename = file_out, plot = a, path = dir_out, device = "pdf")



# BottleRocket2 
custom.col5 <- data.frame(type = levels(data_pca$Edge.angle), col = c("#046C9A", "#FAD510")) data_pca$col <- custom.col5[data_pca$Edge.angle, "col"]


# plots the edge angle  - PCA with convex hull 
b<- autoplot(imp_data.pca, data = data_pca, colour = "Edge.angle", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "convex", frame.colour = "Edge.angle", 
             frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col5$col)
         
print(b)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_angle", ".pdf")
ggsave(filename = file_out, plot = b, path = dir_out, device = "pdf")



# Royal1 
custom.col7 <- data.frame(type = levels(data_pca$Raw.material), col = c("#899DA4", "#DC863B")) 
data_pca$col <- custom.col7[data_pca$Raw.material, "col"]


# plots the raw material  - PCA with convex hull 
c<- autoplot(imp_data.pca, data = data_pca, colour = "Raw.material", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE,
             loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "convex", frame.colour = "Raw.material", 
             frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col7$col)
         
print(c)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_raw.material", ".pdf")
ggsave(filename = file_out, plot = c, path = dir_out, device = "pdf")

```
The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---

# sessionInfo() and RStudio version

```{r}
sessionInfo()

```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
