---
title: "Plots - tool function experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---



```{r Knitr Options, include=FALSE}
	knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)
```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(ggfortify)
library(wesanderson)
```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
```

The imported file is: "`r paste0("~/", data_file)`"  



---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 24:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")
```


---

# Plot each of the selected numeric variable 
## (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
for (i in num.var){
  #plot
  range_var <- range(imp_data[[i]]) # gets the min/max range of the data set 
  
	p <- ggplot(data=imp_data, aes_string(x="Task", y=names(imp_data)[i], colour="Edge.angle")) + 
	     # avoids overplotting
	     geom_jitter(size = 3,  position = position_jitter(width = 0.4, seed = 1)) + 
	     coord_cartesian(ylim = range_var) + 
	     theme_classic() + 
	     # removes the "." between "Edge.angle" in the legend 
	     labs(colour = gsub("\\.", " ", "Edge.angle")) + 
	     # removes the "." between the variable names
	     labs(y = gsub("\\.", " ", names(imp_data)[i])) + 
       # combines the flint and the lydite plot into one    
	     facet_wrap(~Raw.material) 
	print(p)


	#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_plot_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")
}

```


## Principal component analysis 
```{r}
imp_data.pca <- prcomp(imp_data[24:30], scale. = TRUE) 

imp_data.pca$x
pca <- cbind(imp_data[, 1:13], imp_data.pca$x)

# Using ggfortigy
a<- autoplot(imp_data.pca, data = pca, colour = "Edge.angle",
         label = TRUE, label.size = 2,
         loadings = TRUE, loadings.label = TRUE, loadings.label.size  = 3) 

print(a)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_factor_analysis", ".pdf")
ggsave(filename = file_out, plot = a, path = dir_out, device = "pdf")
  

# using ggplot
# PCA use-wear type
b <- ggplot(pca, aes(PC1, PC2, fill = Edge.angle)) +
     stat_ellipse(geom = "polygon", col = "black", alpha = 0.5) +
     scale_fill_manual(values = wes_palette(n = 2, name = "Royal1")) + 
     geom_point(shape = 21, col = "black") +
     labs(fill = "Edge angle") +
     theme_classic()

print(b)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_Edge_angle", ".pdf")
ggsave(filename = file_out, plot = b, path = dir_out, device = "pdf")

# PCA tool type 
c <- ggplot(pca, aes(PC1, PC2,fill = Task)) +
          stat_ellipse(geom = "polygon", col = "black", alpha = 0.5) +
          scale_fill_manual(values = wes_palette(n = 2, name = "Darjeeling2")) +
          geom_point(shape = 21, col = "black") +
          labs(fill = "Task") +
          theme_classic()

print(c)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_Task", ".pdf")
ggsave(filename = file_out, plot = c, path = dir_out, device = "pdf")


```
The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".
---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", full.names = TRUE) %>% 
            md5sum()
```


The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
