---
title: "Plots - tool function experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
editor_options: 
  chunk_output_type: console
---

---

Rbin-file used for this script: **`r paste0("~/", params$data)`**


```{r Knitr Options, include=FALSE}
	knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)
```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- loadObject(paste0(dir_in, "TFE_use-wear.Rbin"))
```

The imported file is: "`r paste0("~/", params$data)`"  



---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 24:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")
```


## Define grouping variable
```{r}
#imp_data[["Sample.Spot"]] <- paste(imp_data[["Sample"]], imp_data[["Location"]], imp_data[["Spot"]], sep="_")
```


---

# Plot each of the selected numeric variable (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
for (i in num.var){
  #plot
	p <- ggplot(data=imp_data, aes_string(x="Task", y=names(imp_data)[i], colour="Edge.angle")) + 
	     geom_jitter(size = 3,  position = position_jitterdodge(jitter.width = 0.5, seed = 123)) + # avoids overplotting
	     theme_classic() + 
	     labs(colour = gsub("\\.", " ", "Edge.angle")) + # removes the "." betwenn "Edge.angle" in the ledgend 
       facet_wrap(~Raw.material) # combines the flint and the lydite plot into one 
	print(p)


	#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_plot_", names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")
}
```
The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".

---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", full.names = TRUE) %>% 
            md5sum()
```


The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
