---
title: "Summary statistics - tool function experiment" 
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
editor_options: 
  chunk_output_type: console
---

Rbin-file used for this script: **`r paste0("~/", params$data)`**


```{r Knitr Options, include=FALSE}
	knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)
```

---


# Goal of the script
This script computes standard descriptive statistics for each group.  
The groups are based on:

* Raw material 
* Spots (replicas) 
* Task + edge anlge


It computes the following statistics:  

* n (sample size = `length`): number of measurements  
* smallest value (`min`)  
* largest value (`max`)
* mean  
* median  
* standard deviation (`sd`)

```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/summary_stats/"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(openxlsx)
library(R.utils)
library(tools)
library(doBy)
```


---

# Get names, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.xlsx$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```


The checksum (MD5 hashes) of the imported file is  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- loadObject(paste0(dir_in, "TFE_use-wear.Rbin"))
```

The imported file is: "`r paste0("~/", params$data)`"  

---


# 4. Define numeric variables
```{r}
num.var <- 24:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")
```

---


# Compute summary statistics
## Create function to compute the statistics at once
```{r}
nminmaxmeanmedsd <- function(x){
	y <- x[!is.na(x)]
	n_test <- length(y)
	min_test <- min(y)
	max_test <- max(y)
	mean_test <- mean(y)
 	med_test <- median(y)
 	sd_test <- sd(y)
 	out <- c(n_test, min_test, max_test, mean_test, med_test, sd_test)
 	names(out) <- c("n", "min", "max", "mean", "median", "sd")
 	return(out)
}
```

##  Compute the summary statistics in groups
### Spots
```{r}
spot <- summaryBy(.~Sample+Location+Area, data=imp_data[c("Sample","Location","Area" ,names(imp_data)[num.var])], FUN=nminmaxmeanmedsd)
str(spot)
```


### Task and edge angel 
```{r}
task <- summaryBy(.~Sample+Task+Edge.angle, data=imp_data[c("Sample", "Task", "Edge.angle",names(imp_data)[num.var])], FUN=nminmaxmeanmedsd)
str(task)
```

### Raw material
```{r}
raw_material <- summaryBy(.~Raw.material+Task+Edge.angle,  data=imp_data[c("Raw.material","Task","Edge.angle" ,names(imp_data)[num.var])], FUN=nminmaxmeanmedsd)
str(raw_material)
```

---

# Save data
## Format name of output file
```{r}
file_out <- "TFE_stats"
```
The file will be saved as "`r paste0("~/", dir_out, ".[ext]")`".


## Write to XLSX
```{r}
write.xlsx(list(spot = spot, task = task, raw_material = raw_material), file = paste0(dir_out, file_out, ".xlsx"))


```


## Save R object
```{r}
saveObject(list(spot=spot, task = task, raw_material = raw_material), file = paste0(dir_out, file_out, ".Rbin"))
```


## Show file information
```{r}
file_out <- c(paste0(dir_out, file_out, ".xlsx"), paste0(dir_out, file_out, ".Rbin"))
md5_out <- md5sum(files_out)
info_out <- data.frame(files = basename(names(md5_out)), checksum = md5_out, row.names = NULL)
```

The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
