---
title: "Plots - tool function experiment"
author: "Lisa Schunk"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: cerulean
    highlight: pygments
    number_sections: true
  github_document: 
    toc: true
    toc_depth: 3
    html_preview: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_format = "all", knit_root_dir = rprojroot::find_rstudio_root_file()) })
---



```{r Knitr Options, include=FALSE}
	knitr::opts_chunk$set(comment=NA, message=FALSE, indent="", error=TRUE)
```

---


# Goal of the script
This script plots all variables to see which ones should be used for further analysis.  
Scatterplot of each variable will be plotted.   



```{r}
dir_in <- "analysis/derived_data/"
dir_out <- "analysis/plots"
```

Raw data must be located in `r paste0("~/", dir_in)`.  
Formatted data will be saved in `r paste0("~/", dir_out)`.

The knit directory for this script is the project directory.
---


# Load packages
```{r}
library(R.utils)
library(ggplot2)
library(tools)
library(tidyverse)
library(ggfortify)

```


---

# Get name, path and information of the file 
```{r}
data_file <- list.files(dir_in, pattern = "\\.Rbin$", full.names = TRUE)
md5_in <- md5sum(data_file)
info_in <- data.frame(file = basename(names(md5_in)), checksum = md5_in, row.names = NULL)
```


The checksum (MD5 hashes) of the imported file is:  
```{r, echo = FALSE}
info_in
```


# Load data into R object
```{r}
imp_data <- loadObject(data_file)
```

The imported file is: "`r paste0("~/", data_file)`"  



---

# Prepare variables
## Define numeric variables
```{r}
num.var <- 24:length(imp_data)
```

The following variables will be used: 

```{r, echo=FALSE}
for (i in num.var) cat("[",i,"] ", names(imp_data)[i], "\n", sep="")
```


---

# Plot each of the selected numeric variable 
## (facet plot = 1 plot for flint, 1 plot for lydite)
```{r}
for (i in num.var){
  #plot
  range_var <- range(imp_data[[i]]) # gets the min/max range of the data set 
  
	p <- ggplot(data=imp_data, aes_string(x="Task", y=names(imp_data)[i], colour="Edge.angle")) + 
	     # avoids overplotting
	     geom_jitter(size = 3,  position = position_jitter(width = 0.4, seed = 1)) + 
	     coord_cartesian(ylim = range_var) + 
	     theme_classic() + 
	     # removes the "." between "Edge.angle" in the legend 
	     labs(colour = gsub("\\.", " ", "Edge.angle")) + 
	     # removes the "." between the variable names
	     labs(y = gsub("\\.", " ", names(imp_data)[i])) + 
       # combines the flint and the lydite plot into one    
	     facet_wrap(~Raw.material) 
	print(p)


	#save to PDF
	file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "_plot_", 
	                   names(imp_data)[i], ".pdf")
	ggsave(filename = file_out, plot = p, path = dir_out, device = "pdf")
}

```


## Principal component analysis 
```{r}
# removes the rows with missing values 
data_pca <- na.omit(imp_data) 

imp_data.pca <- prcomp(data_pca[, c(24:25, 42, 47, 51, 54, 56)], scale. = TRUE) 


#FantasticFox1 = c("#DD8D29", "#E2D200", "#46ACC8", "#E58601", "#B40F20")
custom.col4 <- data.frame(type = levels(data_pca$Task), col = c("#E58601", "#B40F20")) 
data_pca$col <- custom.col4[data_pca$Task, "col"]

# plots the task - PCA with ellipse 
# Using ggfortify
a<- autoplot(imp_data.pca, data = data_pca, colour = "Task", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "norm", frame.colour = "Task", frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col4$col)
         
print(a)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_Task_1", ".pdf")
ggsave(filename = file_out, plot = a, path = dir_out, device = "pdf")


# plots the task - PCA with convex hull
b<- autoplot(imp_data.pca, data = data_pca, colour = "Task", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "convex", frame.colour = "Task", frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col4$col)
         
print(b)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_Task_2", ".pdf")
ggsave(filename = file_out, plot = b, path = dir_out, device = "pdf")


# Chevalier1 = c("#446455", "#FDD262", "#D3DDDC", "#C7B19C")
custom.col5 <- data.frame(type = levels(data_pca$Edge.angle), col = c("#446455", "#FDD262")) 
data_pca$col <- custom.col5[data_pca$Edge.angle, "col"]

# plots the edge angle  - PCA with ellipse 
c<- autoplot(imp_data.pca, data = data_pca, colour = "Edge.angle", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "norm", frame.colour = "Edge.angle", frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col5$col)
         
print(c)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_angle_1", ".pdf")
ggsave(filename = file_out, plot = c, path = dir_out, device = "pdf")


# plots the edge angle  - PCA with convex hull 
d<- autoplot(imp_data.pca, data = data_pca, colour = "Edge.angle", size = 2,
             loadings = TRUE, loadings.colour = "black", loadings.label = TRUE, loadings.label.colour = "black", 
             loadings.label.size  = 4, loadings.label.hjust = 1, loadings.label.vjust = 1,  
             frame = TRUE, frame.type = "convex", frame.colour = "Edge.angle", frame.alpha = 0) + 
             theme_classic() +
             scale_colour_manual(values = custom.col5$col)
         
print(d)

# saves the plot
file_out <- paste0(file_path_sans_ext(info_in[["file"]]), "TFE_PCA_angle_2", ".pdf")
ggsave(filename = file_out, plot = d, path = dir_out, device = "pdf")

```
The files will be saved as "`r paste0("~/", dir_out, ".[ext]")`".
---


# Show plot files information
```{r}
info_out <- list.files(path = dir_out, pattern = "\\.pdf$", full.names = TRUE) %>% 
            md5sum()
```


The checksum (MD5 hashes) of the exported files are:  
```{r, echo = FALSE}
info_out
```


---

# sessionInfo() and RStudio version

```{r}
sessionInfo()
```

RStudio version `r readLines("analysis/scripts/RStudioVersion.txt", n = 1)`.


---

END OF SCRIPT
